A quantum parallel Markov chain Monte Carlo

### 2023-01-18 main.tex
more discrete section

We wish to sample from a target distribution that has probability mass function $\pi(\cdot)$ defined over a discrete set $\{\ttheta_\alpha\}_{\alpha \in \mathcal{A}}$, for $\mathcal{A}$ some finite or countably infinite set of indices.

\subsubsection{Ising model on a 2-dimensional lattice}

@@

Here, we are interested in an Ising-type lattice model over $\ttheta \in \{-1,1\}^D$, where $D$ is the number of spins $\theta_d \in \{-1,1\}$ and $|\mathcal{A}|=2^D$. In particular, we consider targets of the form
\begin{align}\label{eq:ising}
	\pi (\ttheta | \rho) &\propto \exp \left( \rho \sum_{(d,d')\in \mathcal{E}} \theta_d \theta_{d'} \right)  \, ,
\end{align}
where $\rho>0$ is the interaction term and $\mathcal{E}$ is the lattice edge set.  We specify a two-dimensional 500-by-500 lattice with nearest neighbors connections and fix $\beta=1$.  Since this latter setting encourages neighboring spins to equal each other, we begin our QPMCMC trajectories at the lowest-probability `checkerboard' state as initial configuration.  At each iteration, we generate collections of proposal states by uniformly sampling $P$ individual spins $\theta_p$, $p \in \{1,\dots,P\}$, and obtaining each proposal state $\ttheta_p$ by updating the current state $\ttheta_0$ at the single location corresponding to $\theta_p$. Figure \ref{fig:ising2d} shows results from 10 independent runs using $P\in \{4, 8, 16, \dots, 2048\}$ proposals.  Trace plots of unnormalized log-probabilities indicate that higher proposal counts enable discovery of higher probability configurations.  Interestingly, QPMCMC appears to be particularly beneficial in this large $P$ context, requiring less than one-tenth the target evaluations required using conventional computing when $P=2048$.

\subsubsection{Bayesian image something}

>>>>

We wish to sample from a target distribution defined over a discrete set $\{\ttheta_\alpha\}_{\alpha \in \mathcal{A}}$, for $\mathcal{A}$ some finite or countably infinite set of indices.  \citet{glatt} establish the broader measure theoretic foundations for parallel MCMC procedures that use selection probabilities \eqref{eq:probs} to maintain detailed balance. In particular, when the target distribution has probability mass function $\pi(\cdot)$ defined with respect to the counting measure on the power set $2^{\mathcal{A}}$, then detailed balance results in the kernel $P(\cdot,\cdot)$  satisfying
\begin{align*}
	\pi(\alpha) = \sum_{{\alpha'}} \pi(\alpha') P(\alpha',\alpha) \, , \quad \forall \alpha,\alpha' \in \mathcal{A} \, .
\end{align*}
Sections \ref{sec:ising} and \ref{sec:bayesIm} consider targets defined over finite sets $\mathcal{A}$ and make use of the uniform independence proposal scheme described in Section \ref{sec:sym}, thereby facilitating simplified acceptance probabilities \eqref{eq:probsSimp}.  This scheme proposes single-flip updates to the current Markov chain state $\ttheta_0$, but it is worth noting that other multiple-flip schemes would also be amenable to QPMCMC.

\subsubsection{Ising model on a 2-dimensional lattice}\label{sec:ising}

@@

Here, we are interested in an Ising-type lattice model over configurations $\ttheta=(\theta_1,\dots,\theta_D)$ consisting of $D$ individual spins $\theta_d \in \{-1,1\}$. In terms of the preceding section, we have $\mathcal{A}=\{-1,1\}^D$ and $|\mathcal{A}|=2^D$. In particular, we consider targets of the form
\begin{align}\label{eq:ising}
	\pi (\ttheta | \rho) &\propto \exp \left( \rho \sum_{(d,d')\in \mathcal{E}} \theta_d \theta_{d'} \right)  \, ,
\end{align}
where $\rho>0$ is the interaction term and $\mathcal{E}$ is the lattice edge set.  We specify a two-dimensional 500-by-500 lattice with nearest neighbors connections and fix $\rho=1$.  Since this latter setting encourages neighboring spins to equal each other, we begin our QPMCMC trajectories at the lowest-probability `checkerboard' state for an initial configuration.  At each iteration, we generate collections of proposal states by uniformly flipping $P$ individual spins $\theta_p$, $p \in \{1,\dots,P\}$, and obtaining each proposal state $\ttheta_p$ by updating the current state $\ttheta_0$ at the single location corresponding to $\theta_p$. Figure \ref{fig:ising2d} shows results from 10 independent runs using $P\in \{4, 8, 16, \dots, 2048\}$ proposals.  Trace plots of unnormalized log-probabilities indicate that higher proposal counts enable discovery of higher probability configurations.  Interestingly, QPMCMC appears to be particularly beneficial in this large $P$ context, requiring less than one-tenth the target evaluations required using conventional computing when $P=2048$.

\subsubsection{Bayesian image something}\label{sec:bayesIm}
